---
title: "Homework 1"
output: github_document
---
Name: Xi Peng

UNI: xp2213

Date: 09.19.2024

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```


# Problem 1
## Section 1 - Data import
```{r, echo=FALSE, Message=FALSE}
data("penguins", package = "palmerpenguins")
```
The "penguins" dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.
The important variables are: `r names(penguins)`.The mean flipper length `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`
mm.

## Section 2 - Scatterplot of Flipper length vs. Bill Length

```{r, echo=FALSE, Message=FALSE}
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point(na.rm=TRUE)
ggsave("Scatterplot_of_Flipper_Length_and_Bill_Length.pdf", height = 4, width = 6)
```

The scatterplot has been saved as pdf format and named "Scatterplot_of_Flipper_Length_and_Bill_Length.pdf".


# Problem 2
## Section 1 - Data frame
```{r, echo=FALSE, Message=FALSE}
set.seed(1)
problem2samp_df = tibble(
    random_sample = rnorm(10),
    vec_logical = rnorm(10) > 0,
    vec_char = sample(letters, 10),
    vec_factor = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))
)
```
Data frame was created.
```{r echo=FALSE}
problem2samp_df
```

## Section 2 - Mean

```{r, echo=FALSE, Message=FALSE, warning=FALSE, results='hide'}
mean(pull(problem2samp_df, random_sample), na.rm = TRUE)
mean(pull(problem2samp_df, vec_logical), na.rm = TRUE)
mean(pull(problem2samp_df, vec_char), na.rm = TRUE)
mean(pull(problem2samp_df, vec_factor), na.rm = TRUE)
```
In this process, I was able to find means for random_sample and ver_logical, which are `r mean(pull(problem2samp_df, random_sample), na.rm = TRUE)` and `r mean(pull(problem2samp_df, vec_logical), na.rm = TRUE)`, respectively. However, there were warning messages for vec_char and vec_factor, indicating that these two variables are not numeric or logical, and thus returned to NA.

## Section 3 - Using 'as.numeric'

```{r, warning=FALSE}
numeric_logical = as.numeric(pull(problem2samp_df, vec_logical))
numeric_char = as.numeric(pull(problem2samp_df, vec_char))
numeric_factor = as.numeric(pull(problem2samp_df, vec_factor))
```
By using 'as.numeric', the vec_logical and vec_factor were successfully converted to numeric values. However, there was warning message for vec_char, indicating that this variable does not represent numbers, and thus its values were converted to NAs.

By checking each variables that were converted to numeric, I found that the values of numeric_logical are `r numeric_logical`, which consist only 1s and 0s. For numeric_factor, its values are `r numeric_factor`, which correspond to 2, 3, and 1. The variable numeric_char contains NAs.

Retry calculating the mean:
```{r, message=FALSE}
mean(numeric_logical, na.rm = TRUE)
mean(numeric_char, na.rm = TRUE)
mean(numeric_factor, na.rm = TRUE)
```
By trying to calculate the means again, I could find the means for numeric_logical and numeric_factor, which are `r mean(numeric_logical, na.rm = TRUE)` and `r mean(numeric_factor, na.rm = TRUE)`, respectively.

In conclusion, logical variables and factors variables can be converted to numeric or their underlying integer codes, allowing calculating the mean. While character variables are unable to be converted to numeric, the R could not assign numeric values to letters or words. Thus, it is unable to calculate mean for character variables.






