---
title: "Homework 1"
output: github_document
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```


# Problem 1
## Section 1 - Data import
```{r, echo=FALSE, Message=FALSE}
data("penguins", package = "palmerpenguins")
```
The "penguins" dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns.
The important variables are: `r names(penguins)`.The mean flipper length `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`
mm.

## Section 2 - Scatterplot of Flipper length vs. Bill Length

```{r, echo=FALSE, Message=FALSE}
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point(na.rm=TRUE)
ggsave("Scatterplot_of_Flipper_Length_and_Bill_Length.pdf", height = 4, width = 6)
```

The scatterplot has been saved as pdf format and named "Scatterplot_of_Flipper_Length_and_Bill_Length.pdf".


# Problem 2
```{r, echo=FALSE, Message=FALSE}
set.seed(1)
problem2samp_df = tibble(
    random_sample = rnorm(10),
    vec_logical = rnorm(10) > 0,
    vec_char = sample(letters, 10),
    vec_factor = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))
)
```








