---
title: "Homework 1"
output: github_document
---
Name: Xi Peng   UNI: xp2213    Date: 09.21.2024

```{r, echo=FALSE, message=FALSE}

library(tidyverse)
library(dplyr)
library(ggplot2)

```


# Problem 1
## Section 1 - Data import and exploration

```{r}

data("penguins", package = "palmerpenguins")

```

The "penguins" dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns, containing demographic information of `r nrow(penguins)` penguins between `r min(pull(penguins, year))` and `r max(pull(penguins, year))`.
The important variables are: `r names(penguins)`. Specifically, the dataset provides the demographic information about penguins, such as sex, year, bill depth and length, and flipper length, and body mass. Also, island and species are also included. The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`
mm.

```{r, echo=FALSE}

NA_rows = penguins |> 
  select (species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year) |> 
  filter(is.na(species) | is.na(island) | is.na(bill_length_mm) | is.na(bill_depth_mm) | is.na(flipper_length_mm) | is.na(body_mass_g) | is.na(sex) | is.na(year))

```

Based on the inspection of missing values in the dataset, there are 11 penguins with missing values across differnt variables. Specifically, two penguins, one from Togersen Island and one from Biscoe Island, are missing all information. Regarding the variable "sex", 11 penguins are missing this value.


Code for clearing missing values for exploring the minimum, maximum, and mean values:
```{r}

penguins_df = drop_na(penguins)

```

To further explore each variable in detail, there are 3 categories for variable "Island", which are: `r unique(pull(penguins,island))`. The species of sampled penguins are: `r unique(pull(penguins, species))`. After removing the missing values, the minimum, maximum and mean values for bill length are `r min(pull(penguins_df, bill_length_mm))` mm, `r max(pull(penguins_df, bill_length_mm))` mm, and `r mean(pull(penguins_df, bill_length_mm))`, respectively. The minimum, maximum and mean values for bill depth are `r min(pull(penguins_df, bill_depth_mm))` mm, `r max(pull(penguins_df, bill_depth_mm))` mm, and `r mean(pull(penguins_df, bill_depth_mm))`, respectively. The minimum and maximum values for flipper length are `r min(pull(penguins_df, flipper_length_mm))` mm and `r max(pull(penguins_df, flipper_length_mm))` mm, respectively. The minimum, maximum and mean values for body mass are `r min(pull(penguins_df, body_mass_g))` g, `r max(pull(penguins_df, body_mass_g))` g, and `r mean(pull(penguins_df, body_mass_g))`g, respectively.


## Section 2 - Scatterplot of Flipper length vs. Bill Length

```{r, echo=FALSE, Message=FALSE}

ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point(na.rm=TRUE)
ggsave("Scatterplot_of_Flipper_Length_and_Bill_Length.pdf", height = 4, width = 6)

```

The scatterplot has been saved as pdf format and named "Scatterplot_of_Flipper_Length_and_Bill_Length.pdf".


This scatterplot illustrates a clear separation among these three penguin species based on flipper length and bill length, with some overlap in the bill length between Chinstrap and Gentoo penguins. Gentoo penguins have the longest flippers. Chinstrap penguins have shorter flippers but similar bill length to Gentoo penguins. Adelie penguins have the shortest flippers and bills among these three species.


# Problem 2

## Section 1 - Creating a data frame

```{r, Message=FALSE}

set.seed(1)
problem2samp_df = tibble(
    random_sample = rnorm(10),
    vec_logical = rnorm(10) > 0,
    vec_char = c("w","s","e","i","y","o","u","t","r","p"),
    vec_factor = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))
)

```

Data frame was created, as shown below:

```{r echo=FALSE}

problem2samp_df

```

## Section 2 - Initial attempt at calculating the mean of each variable in the created data frame

```{r, echo=FALSE, Message=FALSE, warning=FALSE, results='hide'}

mean(pull(problem2samp_df, random_sample), na.rm = TRUE)
mean(pull(problem2samp_df, vec_logical), na.rm = TRUE)
mean(pull(problem2samp_df, vec_char), na.rm = TRUE)
mean(pull(problem2samp_df, vec_factor), na.rm = TRUE)

```

In this initial attempt, I was able to find means for random_sample and vec_logical, which are `r mean(pull(problem2samp_df, random_sample), na.rm = TRUE)` and `r mean(pull(problem2samp_df, vec_logical), na.rm = TRUE)`, respectively. However, there were warning messages for vec_char and vec_factor, indicating that these two variables are not numeric or logical, and thus returned to NA.

## Section 3 - Exploring type conversion: Using 'as.numeric'

Codes for conversion:

```{r, warning=FALSE, message=FALSE}

numeric_logical = as.numeric(pull(problem2samp_df, vec_logical))
numeric_char = as.numeric(pull(problem2samp_df, vec_char))
numeric_factor = as.numeric(pull(problem2samp_df, vec_factor))

```

By using 'as.numeric', the vec_logical and vec_factor were successfully converted to numeric values. However, there was warning message for vec_char, indicating that this variable does not represent numbers, and thus its values were converted to NAs.

By checking each variables that were converted to numeric, I found that the values of numeric_logical are `r numeric_logical`, which consist only 1s and 0s. For numeric_factor, its values are `r numeric_factor`, which correspond to 2, 3, and 1. The variable numeric_char contains NAs.

Retry calculating the mean:

```{r, Message=FALSE, results='hide'}

mean(numeric_logical, na.rm = TRUE)
mean(numeric_char, na.rm = TRUE)
mean(numeric_factor, na.rm = TRUE)

```

By trying to calculate the means again, I could find the means for numeric_logical and numeric_factor, which are `r mean(numeric_logical, na.rm = TRUE)` and `r mean(numeric_factor, na.rm = TRUE)`, respectively.

In conclusion, logical variables and factors variables can be converted to numeric or their underlying integer codes, allowing calculating the mean. While character variables are unable to be converted to numeric, the R could not assign numeric values to letters or words. Thus, it is unable to calculate mean for character variables.






